---
title: "Modeling home prices using realtor data"
author: "Nimo"
date: "1 March 2016"
output: html_document
---
#### Bullet 4,5 and 6

Reading the data into R, removing the columns and renaming the variables;

```{r,echo=FALSE}
HP<-read.table("http://www.amstat.org/publications/jse/datasets/homes76.dat.txt",header=TRUE)

HP<-HP[-c(1,7,10,15,16,17,18,19)]

colnames(HP)<-c("price","size","lot","bath","bed","year","age","garage","status","active","elem")

```

#### Bullet 7
Preview of the data;

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(DT)
datatable(HP)

```

#### Bullet 8

Performing data exploration;

```{r,echo=FALSE}
str(HP)
summary(HP)
head(HP,10)

```

Price appears to be affected by the following variables: size, lot, bath, bed, age and garage.

#### Bullet 9 and 10

The range for prize in thousands of dollars is;

```{r,echo=FALSE}
range(HP$price)
```

The range for size in thousands of square feet is;

```{r,echo=FALSE}
range(HP$size)

```

Creating models using forward selection and backward elimination;

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(MASS)

mod.fs<-stepAIC(lm(price~1,data=HP),direction="forward",scope=~size+lot+bath+bed+age+garage+active+elem)

mod.be<-stepAIC(lm(price~size+lot+bath+bed+age+garage+active+elem,data=HP),direction="backward")

```

From the results, they both result in the same model. 

#### Bullet 11

Regressing all variables except status and year;

```{r,echo=FALSE}
mod1<-lm(price~size+lot+bath+bed+age+garage+active+elem,data=HP)
summary(mod1)
```

Plotting the residuals of each variable;

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(car)
residualPlots(mod1)

```

Based on the residual plots, a non-linear model is more appropriate. The adjusted R-squared value for the model is 0.4411.

#### Bullet 12

Adding the interaction between bath and bed and the square of age to the model;

```{r,echo=FALSE}

mod2<-update(mod1,~.+bath:bed+I(age^2))
summary(mod2)

```

The R-squared value for this model is 0.5068.

#### Bullet 13

Using only harris and edison from the column elem;
```{r,echo=FALSE}

A<-subset(HP,HP$elem=="harris"|HP$elem=="edison")
mod3<-update(mod2,data=A)
  
summary(mod3)

```

Comparing the last two models;

```{r,echo=FALSE}

Anova(mod3,mod2)
```

#### Bullet 14

Mean square prediction error for the models:
Forward selection model;
```{r,echo=FALSE}

pred1<-predict(mod.fs,type="response")
n1<-length(pred1)
index1<-sample(1:n1,size=0.7*n1)
train1<-pred1[index1]
test1<-pred1[-index1]
mean((mean(train1)-test1)^2)
#mean((mean(test1)-train1)^2)

```

Backward elimination model;
```{r,echo=FALSE}
pred2<-predict(mod.be,type="response")
n2<-length(pred2)
index2<-sample(1:n2,size=0.7*n2)
train2<-pred2[index2]
test2<-pred2[-index2]
mean((mean(train2)-test2)^2)
#mean((mean(test2)-train2)^2)

```

Model with all variables;
```{r,echo=FALSE}
pred3<-predict(mod1,type="response")
n3<-length(pred3)
index3<-sample(1:n3,size=0.7*n3)
train3<-pred3[index3]
test3<-pred3[-index3]
mean((mean(train3)-test3)^2)
#mean((mean(test3)-train3)^2)

```

Model including age^2 and bath:bed;
```{r,echo=FALSE}
pred4<-predict(mod2,type="response")
n4<-length(pred4)
index4<-sample(1:n4,size=0.7*n4)
train4<-pred4[index4]
test4<-pred4[-index4]
mean((mean(train4)-test4)^2)
#mean((mean(test4)-train4)^2)

```

Model using only harris and edison from elem;
```{r,echo=FALSE}
pred5<-predict(mod3,type="response")
n5<-length(pred5)
index5<-sample(1:n5,size=0.7*n5)
train5<-pred5[index5]
test5<-pred5[-index5]
mean((mean(train5)-test5)^2)
#mean((mean(test5)-train5)^2)

```

The model that uses only harris and edison from the variable elem has the smallest training mean squared prediction error. This model also has the smallest test mean squared prediction error.

#### Bullet 15

95% prediction interval for: 1879 square feet, lot category 4, 2.5 baths, 3 bedrooms, built in 1975, 2-car garage and near Parker Elementary. 
```{r,echo=FALSE}
B<-data.frame(size=1.879,lot=4,bath=2.5,bed=3,year=1975,age=0.5,garage=2,active=0:1,elem="parker")


B$pred.price<-predict(mod2,newdata=B,level=0.95,type="response")
B

```

The predicted price for a home with the above specification that is sold is 242,446.60 dollars while one that is still pending is 272,803.80 dollars.

#### Bullet 16

Visualising the effects;

```{r,echo=FALSE,message=FALSE,warning=FALSE}

library(effects)
plot(allEffects(mod2))
```

The graphs show that size, lot, garage and active have a positive correlation with price. For the elem varibale, edison has the highest effect on price. Age and age^2 have decreasing effect on price at first then the effect grows as the value becomes positive. For the interaction term between bath and bed, the effect is decresing at first then it increases as the value goes beyond 2.

```{r,echo=FALSE}
plot(effect("bath*bed", mod2))
```

This graph looks at the effect of the interaction term bath:bed on price. The effect of the predictor is decreasing at lower values then it increases at higher values.

```{r,echo=FALSE}
plot(effect("bath*bed", mod2, xlevels=list(bed=2:5)))
```

For the values of bed that lie between 2 and 5, the effect is still clear as it is decreasing at lower values but increasing at higher values of bath.
 
```{r,echo=FALSE}
plot(effect("bath*bed", mod2, xlevels=list(bath=1:3)))

```

For the values of bath that lie between 1 and 3, the effect is declining at lower values of bed but increasing at higher values.


